The Impact of Urbanization, Culture, and Wealth on Public Art in Toronto

### 2024-01-23 outputs/paper/paper.qmd
Edited the pdf and paper.

---
title: "The Impact of Urbanization, Culture, and Wealth on Public Art in Toronto"
author: 
  - Alaina Hu
thanks: "Code and data are available at: LINK."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true
number-sections: true
# bibliography: references.bib
---

@@

#install.packages("modelsummary")
library(tidyverse)
library(knitr)
library(dplyr)
library(modelsummary)
library(ggplot2)
```

@@

## Public Art
The Public Art raw dataset includes information such as the source, artist, title, medium, installation year, and ward location on the 409 public art pieces
in Toronto. Since we are only interested in the art pieces and their respective
ward location, we focus on these two aspects of the data. Figure 1, shows the part of the public art data we focus on.

```{r}
#| label: cleaned_art
#| fig-cap: Sample of cleaned public art data
#| echo: false
art_data <- read_csv("/cloud/project/outputs/data/cleaned_art_data.csv", show_col_types = FALSE)

art_data |>
  slice(1:5) |>
  kable(col.names = c("Art ID", "Ward", "Ward Name"))

@@

<!-- Talk more about it. -->

>>>>

---
title: "The Impact of Urbanization, Culture, and Wealth on Public Art in Toronto"
author: 
  - Alaina Hu
thanks: "Code and data are available at: https://github.com/alainahu/starter_folder."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
toc: true
number-sections: true
# bibliography: references.bib
---

@@

#install.packages("modelsummary")
#install.packages("sf")

library(tidyverse)
library(knitr)
library(kableExtra)
library(dplyr)
library(modelsummary)
library(ggplot2)
library(sf)

@@

## Public Art
The Public Art raw dataset includes information such as the source, artist, title, medium, installation year, and ward location on the 409 public art pieces
in Toronto. Since we are only interested in the art pieces and their respective
ward location, we focus on these two aspects of the data. Table 1, shows the part of the public art data we focus on.
\vspace{10mm}

```{r}
#| label: cleaned_art
#| tbl-cap: Sample of cleaned public art data
#| echo: false
art_data <- read_csv("/cloud/project/outputs/data/cleaned_art_data.csv", show_col_types = FALSE)

art_data |>
  slice(1:5) |>
  kable(col.names = c("Art ID", "Ward", "Ward Name")) |>
  kable_styling(latex_options = c("hold_position"))

@@

## Ward Profiles (25-Ward Model)

In the 25-Ward model, the city of Toronto is separated into 25 neighborhoods or
voting districts. To better understand the effect of urbanization, cultural diversity, and wealth on the location of public art pieces, we compare the effect
of population, minority population, income on the number of art works across the 
25 wards. The `Ward Profiles (25-Ward Model)` data contains demographic, social, and economic information for each ward such as population, households, families, education, ethnocultural composition, spoken languages, income and housing costs. For our purpose of research, we are interested in population, minority population, and average income of each ward, so we extract these pieces of information. A sample of the cleaned dataset for the wards is shown below in Figure 2. 
\vspace{10mm}

```{r}
#| label: cleaned_ward
#| tbl-cap: Sample of cleaned Toronto 25 Ward data
#| echo: false
ward_data <- read_csv("/cloud/project/outputs/data/cleaned_ward_data.csv", show_col_types = FALSE)

ward_data |>
  slice(1:5) |>
  kable(col.names = c("Ward", "Population", "Minority Population", "Income")) |>
  kable_styling(latex_options = c("hold_position"))

## Public Art by Ward
As we are interested in the number of public art pieces in each ward, we take the cleaned art data and group the wards together. To better understand our dependent response variable of the number of public art installations by ward, we organize the artworks by ward collect a count. Figure 3 shows the number of art pieces by ward for the 25 wards of Toronto. To better visualize the spread of the number of pieces in each ward, we create a bar graph shown in Figure 4.

\newpage

```{r}
#| label: art_by_ward
#| tbl-cap: Number of public art installations by ward.
#| echo: false
art_ward_data <- read_csv("/cloud/project/outputs/data/analysis_data.csv", show_col_types = FALSE)

art_ward_data |>
  select(WARD,ward_name, n) |>
  kable(col.names = c("Ward", "Ward Name", "Number of Public Art Pieces")) |>
  kable_styling(latex_options = c("hold_position"))

\newpage
```{r}
#| label: art_by_ward_bar
#| tbl-cap: Number of public art installations by ward.
#| echo: false
art_ward_data |>
  ggplot(data = art_ward_data, mapping = aes(x = reorder(WARD, n), y = n)) +
  geom_bar(stat = "identity") +
  theme_light() +
  labs(x = "Ward", y = "Number of Public Art Pieces")

```
From Figure 4, we can see that Ward 10: Spadina-Fort York, Ward 13: Toronto-Centre, and Ward 11: University-Rosedale have the most number of public art installations while Ward 7: Humber River-Black Creek, Ward 9: Davenport, and Ward 1: Etobicoke North have the least number of public art installations. This aligns with intuition and literature regarding the connection between public art and urbanization. The wards with the

```{r}
library(sf)
#| echo: false

# URL to the zip file
url <- "https://ckan0.cf.opendata.inter.prod-toronto.ca/dataset/5e7a8234-f805-43ac-820f-03d7c360b588/resource/35f67d86-cfc8-4483-8d77-50d035b010d9/download/25-ward-model-december-2018-wgs84-latitude-longitude.zip"

# Temporary file to store the downloaded zip
temp_zip <- tempfile(fileext = ".zip")

# Download the zip file
download.file(url, temp_zip, mode = "wb")

# Unzip the file to a temporary directory
temp_dir <- tempdir()
unzip(temp_zip, exdir = temp_dir)

# Assuming the shapefile is directly in the unzipped folder and has a standard .shp extension
shapefiles <- list.files(temp_dir, pattern = "\\.shp$", full.names = TRUE)

# Read the shapefile (replace 'shapefiles[1]' with the specific file if there are multiple shapefiles)
toronto_map <- st_read(shapefiles[1])

# Clean up the downloaded zip file
unlink(temp_zip)

ggplot(data = toronto_map) +
  geom_sf() +
  theme_minimal()

```
<!-- Talk more about it. -->
